### 2020-11-10 manuscripts/owp_fim3_2020/sec/keypoints.tex

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item The National Water Model produces forecasts at more than 2.8 million river reaches in the United States.
\item OWP FIM produces flood inundation maps by converting discharges to stages and stages to extents and depths.
\item FIM 3 is a modified version of the Height Above Nearest Drainage technique that provides version improvement with FIM 1 and 2.
\end{keypoints}