### 2021-01-07 manuscripts/owp_fim3_2020/dependencies/trackchanges-0.7.0/testing/testing_new.tex

\documentclass{report}

\usepackage{color} % Allow text colors

%\usepackage{fullpage}

%\usepackage[finalnew]{../LatexPackage/trackchanges}
%\usepackage[finalold]{../LatexPackage/trackchanges}
%\usepackage[footnotes]{../LatexPackage/trackchanges}
%\usepackage[inline]{../LatexPackage/trackchanges}
%\usepackage[margins]{../LatexPackage/trackchanges}
%\usepackage[margins, movemargins]{../LatexPackage/trackchanges}
\usepackage[margins, adjustmargins]{../LatexPackage/trackchanges}

% Setup commands for 6 editors (one more than allowed).
\addeditor{novi}
\addeditor{ym}
\addeditor{three}
\addeditor{quatro}
\addeditor{pyat}
\addeditor{one more}

\title{TrackChanges Testing Document}
\author{Monty Python's Flying Circus}
\date{April, 2009}

\begin{document}
\maketitle

\chapter{Court scene - multiple murderer}

\emph{Cut to a courtroom. Severe atmosphere.}

\noindent\\ Judge:

Michael Norman Randall, you have been found guilty of the murder of Arthur Reginald Webster, Charles Patrick Trumpington, Marcel Agnes Bernstein, Lewis Anona Rudd, John Malcolm Kerr, Nigel Sinclair Robinson, Harry Potter, Thing 1 \& Thing 2, Humpty Dumpty, Superman, Cold Fusion, The Lorax, Mr. T, The Dread Pirate Roberts, Felicity Jayne Stone, Jean-Paul Reynard, Rachel Shirley Donaldson, Stephen Jay Greenblatt, Karl-Heinz Mullet, Belinda Anne Ventham, Juan-Carlos Fernandez, Thor Olaf Stensgaard, Lord Kimberley of Pretoria, The Right Honourable Nigel Warmsly Kimberley, Robert Henry Noonan, Your Mom and Felix James Bennett, on or about the morning of the 19th December 1972. Have you anything to say before I pass sentence?

\noindent\\ Randall:

Yes, sir. I'm very sorry.

\noindent\\ Judge:

Very sorry?

\noindent\\ Randall:

Yes, sir. It was a very very bad thing to have done and I'm really very ashamed of myself. I can only say it won't happen again. To have murdered so many people in such a short space of time is really awful, and I really am very, very, very sorry that I did it, and also that I've taken up so much of the court's valuable time listening to the sordid details of these senseless killings of mine. I would particularly like to say, a very personal and sincere 'sorry' to you, m'lud, for my appalling behaviour throughout this trial. I'd also like to say sorry to the police, for putting them to so much trouble (shot of three heavily bandaged exhausted-looking policemen behind him) for the literally hours of work they've had to put in, collecting evidence and identifying corpses and so forth. You know I think sometimes we ought to realize the difficult and often dangerous work involved in tracking down violent criminals like myself and I'd just like them to know that their fine work is at least appreciated by me.

\noindent\\\emph{The policemen look embarrassed.}

\noindent\\ First Policeman:

No, no, we were only doing our job.

\noindent\\ Second Policeman:

No, no, no, no.

\noindent\\ Randall:

It's very good of you to say that, but I know what you've been through.

\noindent\\ First Policeman:

No, no, we've had worse.

\noindent\\ Third Policeman:

It was plain sailing apart from the arrest.

\noindent\\ Randall:

I know and I'm grateful. I'd like to apologize too to the prosecuting counsel for dragging him in here morning after morning in such lovely weather.

\noindent\\ Randall:

And now I must come to the jury. What can I say. I've dragged you in here, day after day, keeping you away from your homes, your jobs, your loved ones, just to hear the private details of my petty atrocities.

\noindent\\ Foreman:

No, no, it was very fascinating.

\noindent\\ Randall:

But you could have had a much nicer case.

\noindent\\ Foreman:

No, no, murder's much more fun.

\noindent\\ First Juryman:

Yes and so many of them.

\noindent\\ Second Juryman:

Excellent.

\noindent\\ Third Juryman:

We've had a terrific time. (the jury applauds)

\noindent\\ Randall:

(blows his nose, does a Dickie Attenborough) I'm sorry, I'm very moved. And so, m'lud, it only remains for you to pass the most savage sentence on me that the law can provide.

\noindent\\ Judge:

Well er... not necessarily.

\noindent\\ Randall:

No, m'lud, the full penalty of the law is hardly sufficient. I insist I must be made an example of.

\noindent\\ Judge:

Well yes and no. I mean society at large...

\noindent\\ Randall:

Oh no, m'lud. Not with mass murder.

\noindent\\ Judge:

But in this case, (to court) don't you think?

\noindent\\ Court:

Yes, yes!

\noindent\\ Randall:

Oh, come on, m'lud, you've got to give me life.

\noindent\\ Court:

No, no, no, no.

\noindent\\ Randall:

(to court at large) Well, ten years at least.

\noindent\\ Judge:

Ten years!

\noindent\\ Court:

Shame. Shame!

\noindent\\ Randall:

Well five then. Be fair.

\noindent\\ Judge:

No, no. I'm giving you three months.

\noindent\\ Randall:

Oh no, that's so embarrassing. I won't hear of it. Give me six...please.

\noindent\\ Judge:

Well, all right. Six months.

\noindent\\ Randall:

Thank you, m'lud.

\noindent\\ Judge:

But suspended.

\noindent\\  Randall:

Oh no.

\noindent\\ Court:

Hooray. (they applaud)

\noindent\\ Foreman:

Three cheers for the defendant. Hip. Hip.

\noindent\\ Court:

Hooray.

\noindent\\ Foreman:

Hip. Hip.

\noindent\\ Court:

Hooray.

\noindent\\ Foreman:

Hip. Hip.

\noindent\\ Court:

Hooray.

\noindent\\ All:

For he's a jolly good fellow, For he's a jolly good fellow, For he's a jolly good fellow...

\noindent\\ Voice \emph{(off)}:

Which nobody can deny.

\end{document}