### 2021-02-03 manuscripts/owp_fim3_2020/sections/keypoints.tex

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item Office of Water Prediction Flood Inundation Mapping Version 3 `Cahaba' was developed to apply the latest mapping methodology and datasets to produce inundation maps from the United States National Water Model Version 2.1 at continental scale.
\item OWP-FIM V3 `Cahaba' applies advanced techniques and datasets such as the latest hydrography, enhanced computational performance, channel and levee enforcement, two-tiered drainage density, and Manning's n parameter calibration by stream order.  
\item Cahaba was compared to Base Flood Elevation modeled maps and shown to provide version improvement when compared to previous OWP-FIM versions.
\end{keypoints}