### 2021-05-16 manuscripts/owp_fim3_2020/sections/keypoints.tex
more updates

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item Height Above Nearest Drainage (HAND) is a terrain index that detrends elevation data to the nearest relevant drainage path and can be used for rapid generation of flood inundation maps (FIM) across large spatial scales and high spatial resolutions specifically with applications with the United States National Water Model.
\item Despite its advantages, FIM generated from HAND suffers from a number of limitations most notably its dependence on drainage density of the stream network employed.
\item The Office of Water Prediction at the National Oceanic and Atmospheric Administration presents a means of resolving stream density by reducing drainage density to a singular flow path for HAND based FIM. More recent developments in FIM based HAND are also employed in this continental scale model. 
\end{keypoints}