### 2021-05-16 manuscripts/owp_fim3_2020/sections/results.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computational Performance}
\label{ssec:compuational_performance}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
All computational experiments were conducted on an Amazon Web Services (AWS) Elastic Cloud Compute (EC2) T3 instance (2x-large size) with 4 cores, 2 threads per core, and 32GB of random access memory (RAM).
An Elastic Block Storage (EBS) solid state drive (SSD) of the general purpose 2 (gp2) type was used for 2TB of storage.
All tests took place within the project's Docker image providing the operating system and software dependencies.
Table \ref{tab:comp_results} details the computational metrics including CPU time, RAM usage, and data storage footprint for both the hydrofabric production and NRT generation of FIMs.
For the hydrofabric production, computational results were produced for the entire NWM FIM domain noted in \ref{ssec:datasets}.
For the NRT production of FIM from the hydrofabric, computational results were aggregated for the 48 HUC study area.
Computation of the depths and inundation rasters as well as the inundation in polygon form are included in this anaylsis.
Only a subset of NWM FIM domain ever gets mapped for operational purposes to avoid computational burdens by applying a 1.5\% NWM recurrence flow threshold.

\begin{table}
\caption{Descriptive computational statistics for `Cahaba' hydrofabric and NRT FIM production.}
\label{tab:comp_results}
\centering
\begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
%\hline
%"" & \multicolumn{7}{l}{CPU Time (Hrs)} & \multicolumn{7}{l}{RAM (GB)} \\
\hline
Variable & Statistic & Hydrofabric & FIM \\
\hline
\multirow{7}{*}{CPU Time (hrs)} &
n$^{a}$ & XX & XX \\ 
\cline{2-4}
& Sum  & XX & XX \\ 
\cline{2-4}
& Mean & XX & XX \\ 
\cline{2-4}
& Median & XX & XX \\ 
\cline{2-4}
& SD$^{b}$ & XX & XX \\ 
\cline{2-4}
& Max & XX & XX \\ 
\cline{2-4}
& Min & XX & XX \\ 
\hline
\multirow{7}{*}{RAM (GB)} &
n$^{a}$ & XX & XX \\
\cline{2-4}
& Sum  & XX & XX \\
\cline{2-4}
& Mean & XX & XX \\
\cline{2-4}
& Median & XX & XX \\
\cline{2-4}
& SD$^{b}$ & XX & XX \\
\cline{2-4}
& Max & XX & XX \\
\cline{2-4}
& Min & XX & XX \\
\hline
\multirow{7}{*}{Data Storage (GB)} &
n$^{a}$ & XX & XX \\
\cline{2-4}
& Sum  & XX & XX \\
\cline{2-4}
& Mean & XX & XX \\
\cline{2-4}
& Median & XX & XX \\
\cline{2-4}
& Std.Dev.$^{b}$ & XX & XX \\
\cline{2-4}
& Max & XX & XX \\
\cline{2-4}
& Min & XX & XX \\
\hline
\multicolumn{2}{l}{$^{a}$Sample size. $^{b}$Standard Deviation}
\end{tabular}
\end{table}

For comparison purposes, OWP FIM v1 measured that the entire CONUS area excepting a few HUC6's took 1.34 CPU years \cite{liu2016cybergis}. 
OWP FIM v2, developed by \citeA{djokic2019arc}, was estimated to take 0.55 CPU years for the entire NWM stream resolution but lacked the ability to run in parallel on HPC environments.
By utilizing solid programming practices and optimal algorithm selection, we were able to reduce computational requirements while enhancing hydrologic performance.

% TEST SCRIPT: 
% for huc in $(cat /data/temp/ble_test_cases_fa_20210418.lst); do echo HUC: $huc; /usr/bin/time -v /foss_fim/tools/inundation.py -r /data/outputs/first_gms_batch_test/"$huc"/rem_zeroed_masked.tif -c /data/outputs/first_gms_batch_test/"$huc"/gw_catchments_reaches_filtered_addedAttributes.tif -b /data/outputs/first_gms_batch_test/"$huc"/gw_catchments_reaches_filtered_addedAttributes_crosswalked.gpkg -t /data/outputs/first_gms_batch_test/"$huc"/hydroTable.csv -f /data/test_cases/ble_test_cases/validation_data_ble/"$huc"/100yr/ble_huc_"$huc"_flows_100yr.csv -p /data/temp/inundation_py_benchmarking_for_manuscript/inundation_"$huc".gpkg -m filter -i /data/temp/inundation_py_benchmarking_for_manuscript/inundation_"$huc".tif -d /data/temp/inundation_py_benchmarking_for_manuscript/depths_"$huc".tif -u /data/inputs/wbd/WBD_National.gpkg -l WBDHU8 -q | tee /data/temp/inundation_py_benchmarking_for_manuscript/test_1.log; done

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mapping Performance}
\label{ssec:mapping_performance}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%