### 2021-05-16 manuscripts/owp_fim3_2020/sections/title_authors.tex

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3},Trevor Grout\affil{1,3}, Gregory Petrochenkov\affil{4}, Nick Chadwick\affil{1,3}, Zachary Wills\affil{1,3}, Fred Ogden \affil{3}, Roland Viger\affil{4}, Jasmeet Judge\affil{2}}

\affiliation{1}{Lynker Technologies, Leesburg, VA, USA}
\affiliation{2}{Center for Remote Sensing, University of Florida, Gainesville, FL, USA}
\affiliation{3}{National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
\affiliation{4}{Water Mission Area, United States Geological Survey, Boulder, CO, USA}

%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{Fernando Aristizabal}{fernando.aristizabal@noaa.gov}