### 2021-05-16 manuscripts/owp_fim3_2020/sections/discussion.tex
added m&m

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{sec:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Talk about calibration limitations. Cite Boulder and Mike J work.

Talk about cross-walking limitations, FEMA to NWM, NWM to FIM.

Fluvial limitations of HAND.

Limitations of FEMA stream network

Discuss how HAND dependence on Horton-Strahler stream order level and drainage density and how it improves with lower confluence points.