### 2021-05-18 manuscripts/owp_fim3_2020/sections/keypoints.tex

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item Height Above Nearest Drainage (HAND) is a terrain index used for rapid generation of flood inundation maps (FIM) across large scales.
\item Despite its advantages, FIM generated from HAND suffers from a number of limitations most notably its dependence on drainage density.
\item Among other developments, here we present a means of resolving stream density by reducing drainage density to a singular flow path. 
\end{keypoints}