### 2021-10-19 manuscripts/owp_fim3_2020/sections/abstract_plainTextSummary.tex
Updates * first draft complete * version sent out to JJ

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT and PLAIN LANGUAGE SUMMARY
%
% A good Abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.

% The Plain Language Summary should be written for a broad audience,
% including journalists and the science-interested public, that will not have 
% a background in your field.
%
% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
% see http://sharingscience.agu.org/creating-plain-language-summary/)
%
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}
The National Water Model Version 2.1 (NWM V2.1) seeks to expand upon existing systems to provide hourly forecast discharges at four time horizons (1, 3, 10, and 30 days) at over 2.7 million locations throughout the continental United States, Hawaii and Puerto Rico.
Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D fluvial inundation extents used for forecasting purposes to protect life and property.
The Office of Water Prediction Flood Inundation Mapping (OWP-FIM) model `Cahaba' is an enhancemant of the Height Above Nearest Drainage (HAND) method developed to improve upon some of HAND's limitations.
Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, enforcement of levee data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and simpilified Manning's N variation.
Most importantly we detail an issue previously mentioned with HAND as its FIM skill is dependent on drainage threshold.
Since our modeling assumptions our different, we illustrate the enhancement of HAND based FIMs skill as the Horton-Strahler stream orders of the stream networks of the HAND processing units are reduced all the way down to unary.
This is accomplished by deriving level paths for the National Water Model stream network and utilizing those definitions for deriving HAND datasets independetly from other level paths.
`Cahaba' was evaluated at 48 Base Level Engineering (BLE) sites furnished by the Federal Emergency Management Agency (FEMA).
The modeled discharges from the BLE sections were used for the 100 and 500 year flood events as input to OWP FIM instead of NWM discharges to isolate out factors related to hydro-meterology.
The resulting maps are compared to the modeled extents from the BLE and binary contigency statistics are computed.
Across the 49 sites, Critical Success Index (CSI) was found to increase across FIM versions driven by improvements to the probability of detection (POD) and slight reduction in False Alarm Ratio (FAR).
These capabalities were implemented in an open-source package available on GitHub (\url{https://github.com/NOAA-OWP/cahaba}) for continued use and collaboration with the research community.
%
\end{abstract}
%
\section*{Plain Language Summary}
Flooding is one of the most impactful natural disasters on life and property.
The United States National Water Model (NWM) seeks to provide flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
In order to convert forecasts flow rates from the NWM to inundation extents, a model is used that converts elevations relative to sea-level to elevations relative to the nearest river bottom elevation.
This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance where rivers converge.
The maps were compared to other, more rigorous models and shown to perform well when compared to previous versions given certain enhancements to the way HAND to computed.
%