### 2022-01-14 manuscripts/owp_fim4_2021/sections/keypoints.tex

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}
%
\begin{keypoints}
\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to limitations at river junctions.
\item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation.
\end{keypoints}
%